Almost compatible functions and infinite length games

### 2015-12-08 articles/almost-compatible-functions-and-topological-games/article.tex
updated proof from Alan

  \begin{theorem}
    There exists a model of \(ZFC\) for which \(\neg\alcompSP{\aleph_\omega}\).
  \end{theorem}

@@

          The interested reader will want to know that it is shown in
          \cite{MR1045371} that if $\kappa$ is a 2-huge cardinal and $j$ is the 2-huge
          embedding with critical point $\kappa$, then with
           $\lambda = j(\kappa)$ one has that
          $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
          (\kappa^{+\omega+1},\kappa^{+\omega})$ holds.
          \bigskip

          Let $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$
          be a scale in $\Pi\{ \lambda^{+n+1} : n\in \omega\}$
           ordered by the usual mod finite coordinatewise
          ordering. For convenience we may assume that $h_\xi(n) \geq
          \lambda^{+n}$ for all $\xi$ and all $n$.  If $P$ is any poset of
          cardinality less than $\lambda^{+\omega}$, then in the forcing
          extension by $P$,  the sequence
          $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$  remains cofinal in
           $\Pi\{ \lambda^{+n+1} : n\in \omega\}$.

@@

          The forcing notion $\mathbb P_0$ is simply the finite condition
          collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
           \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
           $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
           from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
           forcing extension we let $\mathbb P_1$ be the countable condition
           Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
           less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
           $\mathbb P_1$ has cardinality $\lambda$. We treat
          $\mathbb P_1$ as containing $\mathbb P_0$
          as a subposet by identifying each $(p_0,1)$ with $p_0$.
          After forcing with $\mathbb P_0*\mathbb P_1$ we will have
          that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
           $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
           ordinal $\left(\lambda^{+\omega}\right)^V$.

@@

@@

          Let $\{ x_\xi : \xi\in \lambda^{+\omega+1}\}$ be any enumeration of
           $H(\lambda^{+\omega+1})$  such that
           $\{ x_\xi : \xi\in \lambda^{+\omega}\}$ is also equal to
           $H(\lambda^{+\omega})$. We choose this enumeration in such a way that
           $x_\xi\in x_\eta$ implies $\xi<\eta$.
          We use relation symbol $R_0$ to code
          (and well order)
           $(H(\lambda^{+\omega+1}), \in)$ as follows: $(\xi,\eta)\in R_0$
          if and only if $x_\xi\in x_\eta$.
          Let $R_1$ be a binary relation on $\kappa^{+\omega}$ so that
          $(\kappa^{+\omega},R_1)$ is isomorphic to $\mathbb P_0$. Let
           $R_2$ be a binary relation on $\lambda$ so that
           $R_2\cap (\kappa^{+\omega}\times \kappa^{+\omega})=R_1$ and
           $(\lambda,R_2)$ is isomorphic to $\mathbb P_0*\mathbb P_1$.
          Let $\psi$ be the poset isomorphism from $\lambda$ to
           $\mathbb P_0*\mathbb P_1$.

@@

          \begin{proof}[Proof of Claim]
          Suppose  our originally supplied $N$ fails the conclusion of the
          claim.
          We   know that $\kappa^{+\omega}\in N$,
          (  via $R_1$)
          in which case so is $\kappa^{+\omega+1}$.

          Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
          Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
          (in that we have to formalize the enumeration of formulas)
          let $\{\varphi_n : n\in \omega\}$ is the enumeration of all formulas
          in the language $L$, and let  $\ell_n$ be the minimal integer such
          that the free variables of $\varphi_n$ are among
           $\{ v_0, \ldots, v_{\ell_n}\}$.
          Then,   for each tuple
          $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
          of elements of $\lambda^{+\omega+1}$,
           we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
           the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
           $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
          If there is no such $\xi_0$,
             in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
          \ldots,\xi_{\ell_n})$, then set
           $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
          Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
           $X$
          of $N\cup \beta_0$ that satisfies that
           $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
           simply a large algebra, we can generate all the terms $t$ of
          the algebraic operations $\{ f_n : n\in \omega\}$.
          It is easily seen that for each $\zeta \in X$, there is a term
          $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
           for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
           each $\delta_i\in  N\cup \beta_0$.
           Assume that
            $\zeta\in \kappa^{+\omega+1}$. By
            re-indexing the variables in the term we can assume
            that there is an $n\leq m$ so that
             $\delta_i  <\beta_0$ for $1\leq i\leq n$
              and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
              Let $\vec a $ denote
              the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
            Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
           $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
           in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
           less naturally.
           Consider the set
             $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
          \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
           $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
           Now define $s_1(\eta,\vec a) $ to be
           $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
           and choose the
           unique $\zeta_1\in \lambda^{+\omega+1}$ such
           that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
           Note that $\alpha R_0\zeta_1$ holds if and only if
            $ \alpha \in s_0(\eta,\vec a)$, and therefore
           $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
           (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
           t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
           By elementarity then we have that $\zeta_1\in N$,
           and by similar reasoning the supremum, $\zeta_0$,
          of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
             This of course means that $\zeta < \xi_0$.
          \end{proof}

@@

           Since  $\mathfrak c < \kappa^{+\omega+1}$ there
            is a function $\langle \varrho_n : n\in \omega\rangle$
            in $\Pi_n \lambda^{+\omega}   $ such that the sequence
             $\{ h_\xi : \xi \in N\}$ is unbounded mod finite
             in $\Pi_n \varrho_n$ (by Shelah's pcf theory).
             This is in Jech somewhere.
             For each $n$, $\rho_n \leq \sup(N\cap \lambda^{+n+2})$.

             Since $\mathbb P_0$ has cardinality less than $|N|=\kappa^{+\omega+1}$,
              the sequence $\{ h_\xi : \xi \in \Lambda\cap N\}$ remains unbounded
              mod finite in $\Pi_n \varrho_n$ (and in
               $\Pi_n (\varrho_n\cap N)$).  Now pass to the extension by
                $G\cap \mathbb P_0$ and let $H$ be the function
                 $\operatorname{val}_{G}(\dot H)$, and we recall
                 that $f_{\zeta_\xi}(h_\xi(n)) = H(n)$ for all $n\in \omega$.
           Now pass to the full extension $V[G]$ and again, since
            $\mathbb P_1$ was forced to be countably closed,
             the family $\{ h_\xi : \xi \in \Lambda\cap N\}$ is still
             unbounded in $\Pi_n (\varrho_n \cap N)$. We let
             $A$ be the countable set $N\cap \lambda^{+\omega}$,
             and for  each
             $\xi\in \Lambda\cap N$, there is an $n_\xi$ such that
              $f_{\xi}(h_\xi(m)) = f_A(h_\xi(m))$ for all $m>n_\xi$.
              There is a single $n$ so that $\Lambda_n
               = \{\xi\in \Lambda\cap N       : n_\xi = n\}$ has cardinality
                $\omega_1 $, and thus
                $\{ h_\xi : \xi\in \Lambda_n\cap N\}$ is also unbounded
                in $\Pi_n (\rho_n\cap N)$.
          This certainly implies that there is an $m>n$
          such that $\{ h_\xi (m) : \xi\in \Lambda_n\cap N\}$ is infinite.
          This completes the proof since  $f_{A}(h_\xi(m)) = H(m)$
          for all $\xi\in \Lambda_n\cap N$.
  \end{proof}

>>>>

\begin{theorem} It follows from the existence of a $2$-huge cardinal
that there is a model of $ZFC$ for which $\lnot \mathcal A'(\aleph_{\omega})$.
\end{theorem}

@@

            The interested reader will want to know that it is shown in
            \cite{MR1045371} that if $\kappa$ is a 2-huge cardinal
            and $j$ is the 2-huge
            embedding with critical point $\kappa$, then with
             $\lambda = j(\kappa)$ one has that
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$ holds. There is no loss of
            generality to  also assume that GCH holds in this model.

@@

            Let $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$
            be a scale in $\Pi\{ \lambda^{+n+1} : n\in \omega\}$
             ordered by the usual mod finite coordinatewise
            ordering. For convenience we may assume that $h_\xi(n) \geq
            \lambda^{+n}$ for all $\xi$ and all $n$. For each integer
             $m$ the cofinality of the mod finite ordering on
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$
            is the same as it is for the entire product
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$. If $P$
            is any poset of cardinality less than
             $\lambda^{+m}$ then, in the forcing extension by $P$,
            every function in
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$ is bounded
            above by a ground model function.
            It therefore
            follows easily that  in the forcing
            extension by $P$,  the sequence
            $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$  remains cofinal in
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$.

@@

            The forcing notion $\mathbb P_0$ is simply the finite condition
            collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
             \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
             $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
             from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
             forcing extension we let $\mathbb P_1$ be the countable condition
             Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
             less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
             $\mathbb P_1$ has cardinality $\lambda$. We treat
            $\mathbb P*\mathbb P_1$ as containing $\mathbb P_0$
            as a subposet by identifying each $(p_0,1)$ with $p_0$.
            After forcing with $\mathbb P_0*\mathbb P_1$ we will have
            that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
             $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
             ordinal $\left(\lambda^{+\omega}\right)^V$.

            \newpage

@@

            Let $\{ x_\xi : \xi\in \lambda^{+\omega+1}\}$ be any enumeration of
             $H(\lambda^{+\omega+1})$  such that
             $\{ x_\xi : \xi\in \lambda^{+\omega}\}$ is also equal to
             $H(\lambda^{+\omega})$. We choose this enumeration in such a way that
             $x_\xi\in x_\eta$ implies $\xi<\eta$.
            We use relation symbol $R_0$ to code
            (and well order)
             $(H(\lambda^{+\omega+1}), \in)$ as follows: $(\xi,\eta)\in R_0$
            if and only if $x_\xi\in x_\eta$.
            Let $R_1$ be a binary relation on $\kappa^{+\omega}$ so that
            $(\kappa^{+\omega},R_1)$ is isomorphic to $\mathbb P_0$. Let
             $R_2$ be a binary relation on $\lambda$ so that
             $R_2\cap (\kappa^{+\omega}\times \kappa^{+\omega})=R_1$ and
             $(\lambda,R_2)$ is isomorphic to $\mathbb P_0*\mathbb P_1$.
            Let $\psi$ be the poset isomorphism from $(\lambda,R_2)$ to
             $\mathbb P_0*\mathbb P_1$.

@@

            \bgroup
            \def\proofname{Proof of Claim:\/}

            \begin{proof}
            Suppose  our originally supplied $N$ fails the conclusion of the
            claim.
            We   know that $\kappa^{+\omega}\in N$,
            (  via $R_1$)
            in which case so is $\kappa^{+\omega+1}$.

            Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
            Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
            (in that we have to formalize the enumeration of formulas as per
            G\"odel coding)
            let $\{\varphi_n : n\in \omega\}$ be an enumeration of all formulas
            in the language $L$, and let  $\ell_n$ be the minimal integer such
            that the free variables of $\varphi_n$ are among
             $\{ v_0, \ldots, v_{\ell_n}\}$.
            Then,   for each tuple
            $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
            of elements of $\lambda^{+\omega+1}$,
             we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
             the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
             $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
            If there is no such $\xi_0$,
               in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
            \ldots,\xi_{\ell_n})$, then set
             $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
            Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
             $X$
            of $N\cup \beta_0$ that satisfies that
             $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
             simply a large algebra, we can generate all the terms $t$ of
            the algebraic operations $\{ f_n : n\in \omega\}$.
            It is easily seen that for each $\zeta \in X$, there is a term
            $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
             for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
             each $\delta_i\in  N\cup \beta_0$.
             Assume that
              $\zeta\in \kappa^{+\omega+1}$. By
              re-indexing the variables in the term we can assume
              that there is an $n\leq m$ so that
               $\delta_i  <\beta_0$ for $1\leq i\leq n$
                and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
                Let $\vec a $ denote
                the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
              Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
             $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
             in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
             less naturally.
             Consider the set
               $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
            \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
             $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
             Now define $s_1(\eta,\vec a) $ to be
             $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
             and choose the
             unique $\zeta_1\in \lambda^{+\omega+1}$ such
             that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
             Note that $\alpha R_0\zeta_1$ holds if and only if
              $ \alpha \in s_0(\eta,\vec a)$, and therefore
             $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
             (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
             t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
             By elementarity then we have that $\zeta_1\in N$,
             and by similar reasoning the supremum, $\zeta_0$,
            of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
               This of course means that $\zeta < \xi_0$.
            \end{proof}

            \egroup

@@

            Now we need a result from Shelah's pcf theory which
            is proven in  Jech \cite[24.9]{MR1940513}.
             Since  $\aleph_1=\mathfrak c < \kappa^{+\omega+1}$ there
              is a function $\langle \varrho_n : n\in \omega\rangle$
              in $\Pi_n \lambda^{+\omega}   $ such that the sequence
               $\{ h_\xi : \xi \in N\}$ is unbounded mod finite
               in $\Pi_n \varrho_n$
               For each $n$, $\varrho_n \leq \sup(N\cap \lambda^{+n+2})$.
                  Since $\mathbb P_0$ has cardinality $\kappa^{+\omega}$,
            and so less than $|N|=\kappa^{+\omega+1}$, a standard argument
            (analogous to the fact that adding a Cohen real does
             not add a dominating real)
             shows that    the sequence $\{ h_\xi : \xi \in \Lambda\cap N\}$ remains
             unbounded
                mod finite in $\Pi_n \varrho_n$ (and in
                 $\Pi_n (\varrho_n\cap N)$).

            Now pass to the extension by
                  $G\cap \mathbb P_0$ and let $H$ be the function
                   $\operatorname{val}_{G}(\dot H)$, and we recall
                   that $f_{\zeta_\xi}(h_\xi(n)) = H(n)$ for all $n\in \omega$
            and $\xi\in \Lambda$.
             Now pass to the full extension $V[G]$ and again, since
              $\mathbb P_1$ was forced to be countably closed,
               the family $\{ h_\xi : \xi \in \Lambda\cap N\}$ is still
               unbounded in $\Pi_n (\varrho_n \cap N)$ (no new elements
            were added). We let
               $A$ be the countable set $N\cap \lambda^{+\omega}$,
               and for  each
               $\xi\in \Lambda\cap N$, there is an $n_\xi$ such that
                $f_{\xi}(h_\xi(m)) = f_A(h_\xi(m))$ for all $m>n_\xi$.
                There is a single $n$ so that $\Lambda_n
                 = \{\xi\in \Lambda\cap N       : n_\xi = n\}$ has cardinality
                  $\omega_1 $, and thus
                  $\{ h_\xi : \xi\in \Lambda_n\cap N\}$ is also unbounded
                  in $\Pi_n (\rho_n\cap N)$.
            This certainly implies that there is an $m>n$
            such that $\{ h_\xi (m) : \xi\in \Lambda_n\cap N\}$ is infinite.
            This completes the proof since  $f_{A}(h_\xi(m)) = H(m)$
            for all $\xi\in \Lambda_n\cap N$.
            \end{proof}
